---
title: "The Development Gap in Economic Rationality of Future Elites"
author: "Erik Ø. Sørensen"
date: "`r format(Sys.time(), '%d %B, %Y')`"
always_allow_html: true
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
extrafont::loadfonts()
library("multcomp")
library("tidyverse")
library("haven")
library("modelsummary")
library("kableExtra")
library("here")
library("gt")
library("targets")
source("R/functions.R")
```


# Reading data
```{r}
tar_load(df)
tar_load(studysubjects_categorized)
```


# Summaries of background data

## Experiment data

```{r}
df %>% group_by(country) %>%
  summarize(n = n(),
            one_parent_college = mean(one_parent_college, na.rm=TRUE),
            female = mean(sex==2, na.rm=TRUE),
            age = mean(age, na.rm=TRUE),
            worried = mean(yworried, na.rm=TRUE),
            happiness = mean(happy, na.rm=TRUE),
            government_preference = mean(government_preference=="Yes", na.rm=TRUE)) |>
  gt::gt() |>
  gt::fmt_number(columns = c(one_parent_college, female, worried, happiness, government_preference), decimals = 3) |>
  gt::fmt_number(columns = c(happiness), decimals = 2) |>
  gt::fmt_number(columns = age, decimals=1) |>
  gt::fmt_number(columns=n, decimals=0)
```

## Follow up survey in Tanzania

```{r}
tar_load(tanzaniasurvey)
tz_df <- tanzaniasurvey
```


### Where are they from? 
There are 32 **regions** in Tanzania (three of them on Zanzibar). We have 
manually classified self-reported (free-form) birthplace to official region names.

```{r}
tz_df %>% group_by(birth_region) %>%
  summarize( n = n()) %>%
  arrange(-n) %>% 
  mutate( proportion = n/sum(n), 
          ranking = row_number()) %>%
  gt::gt() |> 
  gt::fmt_number(columns = c(n, ranking), decimals = 0) |>
  gt::fmt_number(columns = c(proportion), decimals = 3)
```


Note that the one "Zanzibar" response is not allocated to a proper region, as
no more information than "Zanzibar" was given by the respondent. 

While the participants at lab experiments at major American universities
have been extensively studied, our sample at the University of Dar-es-Salaam
comes from a less studied population, and not from a permanent pool
of lab experiments. We implemented a small survey of our Tanzanian
lab participants (on 215 out of our 216 lab participants). 

While the university is located in Dar-es-Salaam, we have participants
from 25 out of the 30 Tanzanian regions (as of 2012). Those born in 
Dar-es-Salaam are well represented (16.7 %), but there are slightly more 
from Kagera (17.2%) and almost as many from Kilimanjaro (13.5 %). There 
are 6% or less from each of the other regions.



### Regional (PPP) GDP / capita
The regional GDPs per capita  are reported in Table 22 of 

- [NATIONAL ACCOUNTS STATISTICS OF TANZANIA MAINLAND 2012-2018. First Edition IN THE REVISED GDP SERIES BASE YEAR 2015](https://www.nbs.go.tz/nbs/takwimu/na/National_Accounts_Statistics_of_Tanzania_Mainland_2018.pdf). 


Punching the 2012 GDP per capita at current prices

```{r}
TZ_GDP <- tribble(~region, ~gdp2012,
                  "Dodoma", 927536,
                  "Arusha",1754637,
                  "Kilimanjaro", 1725950,
                  "Tanga", 1431271,
                  "Morogoro", 1373059,
                  "Pwani", 1048866,
                  "Dar es salaam", 2419087,
                  "Lindi", 1361215,
                  "Mtwara", 1298007,
                  "Ruvuma", 1726244,
                  "Iringa", 1991561,
                  "Mbeya", 1687330,
                  "Singida", 872866,
                  "Tabora", 1074905,
                  "Rukwa", 1358997,
                  "Kigoma", 848712,
                  "Shinyanga", 1124625,
                  "Kagera", 869527,
                  "Mwanza", 1450639,
                  "Mara", 1319264,
                  "Manyara", 1461955)
```

For all of the mainland, the number was 
```{r}
gdp_tz_mainland <- 1429327
```

The [GDP per capita, PPP (constant 2017 international $) for Tanzania, according to the World Bank](https://data.worldbank.org/indicator/NY.GDP.PCAP.PP.KD?locations=TZ) was 
```{r}
gdp_tz_wbPPP <- 2094
```

Using the national numbers to compute a national correction factor, we get the PPP adjusted GDP/capita 2012 
in 2017 international dollars:

```{r}
TZ_GDP |> mutate(gdp2012ppp = (gdp2012 / gdp_tz_mainland) * gdp_tz_wbPPP) |>
  arrange(-gdp2012ppp) |>
  mutate(rank = row_number()) |>
  dplyr::select(-gdp2012) |>
  gt::gt() |> 
  gt::fmt_number(columns=gdp2012ppp, decimals = 0, use_seps = TRUE)
```







### Field of study
The most popular reported field of study among our respondents is one or more of the social sciences (42.8%),
with science and engineering close behind (31.6%). Education (9.8%) and humanities (8.4%) are also represented,
but there are less than four percent each from business management and law. On average they report
to be `r mean(tz_df$yrs_study, na.rm=TRUE)` years into their study, with `r max(tz_df$yrs_study, na.rm=TRUE)` as the maximum. 

For a more detailed look at the subjects they studied, let's count the number that in some way
reported to have economics as part of their degree. 

```{r message=FALSE, warning=FALSE}
cat_table <- studysubjects_categorized |>
  mutate(contains_economic = str_detect(studysubject, regex("economic", ignore_case = TRUE))) |>
  group_by(contains_economic) |>
  group_by(category) |>
  summarize(n  = sum(n_responses),
            cond_prop_econ = mean(contains_economic)) |>
  mutate(proportion = n/sum(n)) |>
  arrange(-proportion) |>  
  dplyr::select(c(category, proportion, cond_prop_econ)) |>
  gt::gt() |>
  gt::fmt_number(columns=c(proportion, cond_prop_econ), decimals = 3) 
cat_table
cat_table |> as_latex() |> 
  as.character() |> 
  writeLines(con = here::here("tables","studysubjects.tex"))
```



### Financing of expenses

The median reported yearly expenses was `r format(median(tz_df$year_expenses2, na.rm=TRUE), scientific=FALSE)` TZS, corresponding to about 
`r 0.0006 * median(tz_df$year_expenses2, na.rm=TRUE)` USD at 
2012 exchange rates. 

For a fuller analysis of sources of finance, with all non-reported category observations set to zero:

```{r warning=FALSE}
income_table <- tanzaniasurvey |> 
  dplyr::select(c(survey_id, ann_finance1, ann_finance2, ann_finance3, ann_finance4,
                  ann_finance5, ann_finance6, ann_finance7, ann_finance8)) |>
  pivot_longer(starts_with("ann_finance")) |>
  mutate(amount = ifelse(!is.na(value),value,0),
         name = factor(name),
         source = fct_recode(name,
                             "Government support/loan" = "ann_finance1",
                             "Support from family" = "ann_finance2",
                             "Loan from family" = "ann_finance3",
                             "Own work" = "ann_finance4",
                             "Scholarship grant" = "ann_finance5",
                             "Other grants" = "ann_finance6",
                             "Support from NGO" = "ann_finance7",
                             "Other" = "ann_finance8"),
         nonzero = as.numeric(amount>0)) |>
  group_by(survey_id) |> 
  mutate( share = amount / sum(amount)) |>
  group_by(source) |>
  summarize(share_with_positive = mean(nonzero),
            mean_share = mean(share, na.rm=TRUE)) |>
  arrange(-share_with_positive) |>
  dplyr::select(c(source, share_with_positive, mean_share)) |>
  gt::gt() |> gt::fmt_number(decimals=3)
income_table
income_table |> 
  as_latex() |> 
  as.character() |> 
  writeLines(con=here::here("tables","budgetshares.tex"))
```

We also have a couple of more specific questions about support:

```{r}
tz_df |> summarize(any_familysupport = mean(familysupport),
                   mean_familysupport = mean(familysupport_value, na.rm=TRUE),
                   median_familysupport = median(familysupport_value, na.rm=TRUE)) |>
  gt::gt() |>
  gt::fmt_number(columns = any_familysupport, decimals = 3) |>
  gt::fmt_number(columns = c(mean_familysupport,median_familysupport), decimals=0)
```



And for government loans

```{r}
tz_df |> 
  summarize(mean_govloan = mean(govloan>0, na.rm=TRUE),
            mean_govloanshare = mean(govloan, na.rm=TRUE)) |>
  gt::gt() |> gt::fmt_number(decimals=3)
```




# Descriptives of behavioral outcomes

## Means of the main outcomes

Overall means for a striking first figure of the rationalizability scores. First, calculate the means and standard errors 
for both the original and the standardized scores.

```{r}
tar_load(fig_dta)
```

What are the levels in each country and the level differences: 
```{r} 
fig_dta |> dplyr::select(country, outcome_nm, mean) |> 
  pivot_wider(id_cols=outcome_nm, names_from = country, values_from = mean) |>
  mutate(difference = `United States` - Tanzania) |>
  gt::gt() |>
  gt::fmt_number(columns=c(Tanzania, `United States`), decimals=2) |>
  gt::fmt_number(columns=difference, decimals=3) 
```

For how many is the difference between $e^{**}$ and $e^{***}$ less than 0.01? 

```{r}
df |> 
  mutate(smaller_than1p = abs(EU_PQR-FOSD_PQR)<0.01) |>
  group_by(country) |>
  summarize(mean_smaller_than1p = mean(smaller_than1p)) |>
  gt::gt() |> fmt_number(decimals = 3)
```

How many have rationalizability scores above given thresholds?

```{r}
df |> pivot_longer( cols=c(CCEI_PQR,FOSD_PQR,EU_PQR), names_to = "RPstat",values_to="e")  |>
  dplyr::select(country, RPstat, e) |>
  mutate(above90 = e>0.9,
         above95 = e>0.95,
         RPstat = factor(RPstat, levels=c("CCEI_PQR","FOSD_PQR", "EU_PQR"))) |>
  group_by(country, RPstat) |>
  summarize(share_above90 = mean(above90),
            share_above95 = mean(above95)) |>
  arrange(country, RPstat) |>
  gt::gt() |> fmt_number(decimals = 3)
  

```



And what are the correlations between IQ and the rationalizability measures in the US and Tanzania?

```{r}
df |> group_by(country) |> 
  summarize(cor_IQ_CCEI = cor(iq, CCEI_PQR),
            cor_IQ_FOSD = cor(iq, FOSD_PQR),
            cor_IQ_EU= cor(iq, EU_PQR)) |>
  gt::gt() |> 
  gt::fmt_number(decimals=3)
```




Now, the first plot with the rationalizability scores:
```{r}
fig_dta %>% filter(outcome_nm!="iq") %>%
  ggplot(aes(x=outcome_nm, 
             y=mean, 
             ymin = mean - 1.96 * se_mean, 
             ymax = mean + 1.96 * se_mean, 
             fill=country)) + 
  geom_bar(orientation="x", stat="identity", position=position_dodge2()) +
  geom_errorbar(position=position_dodge(0.9), width=0.3) +
  labs(y="Mean \u00B1 95% C.I.", x= element_blank(), fill = "Country") +
  theme_minimal() + 
  ylim(0, 1) 
ggsave(here("graphs", "mean_differences.pdf"), width=16, height=10, units="cm")
```

Similar graph, but with standardized outcomes and including the IQ scores. We decided to change this
graph into differences between countries: I make it differences between US and Tanzania.

```{r}
fig_dtaw <- fig_dta |> 
  pivot_wider(id_cols = outcome_nm, names_from = country, values_from = c(zmean, se_zmean)) |>
  mutate( zdiff = `zmean_United States` - zmean_Tanzania,
          zdiff_se =  sqrt((`se_zmean_United States`)^2 + se_zmean_Tanzania^2 ))
```


```{r}
fig_dtaw %>%
  ggplot(aes(x=outcome_nm, 
             y=zdiff, 
             ymin = zdiff - 1.96 * zdiff_se, 
             ymax = zdiff + 1.96 * zdiff_se)) + 
  geom_col() +
  geom_errorbar(width=0.3) +
  labs(y="Mean difference in standardized outcomes\nUnited States vs Tanzania \u00B1 95% C.I.", x= element_blank()) +
  theme_minimal() 
ggsave(here("graphs", "zmean_differences.pdf"), width=16, height=10, units="cm")
```

In the final figure, each of the four outcomes have been standardized to an overall average of zero and unit variance before
calculating the averages by country.

The exact numbers in the graph are:
```{r} 
fig_dtaw |> gt::gt() |>
  fmt_number(decimals = 2)
```



## Survival function plot 
```{r}
tar_load(cdf_data)
tar_load(survival_gg)
survival_gg 
ggsave(here::here("graphs","survival_graph.pdf"), 
       width=16, height=10, units="cm")
```



# Risk aversion

For Tanzania, we are interested in whether high and low stakes individuals
exhibited different risk aversion. We operationalize risk aversion by
the average fraction of tokens in the cheaper security (so a high number means
low risk aversion).


Now, what are the average risk aversion in the three groups?
```{r}
df |> group_by(country_high) |> 
  summarize(mean = mean(mean_cheapest_asset),
            std.dev = sd(mean_cheapest_asset),
            se = std.dev/sqrt(n())) |>
  gt::gt() |>
  gt::fmt_number(columns=c(mean, std.dev, se), decimals=3)
```



Testing (t-test) for equality between the treatments in Tanzania?
```{r}
df %>% filter(Tanzania==1) %>%
  t.test(mean_cheapest_asset ~ Treatment, data=.)
```

Testing (t-test) for equality between Tanzania low and the United States:

```{r}

df %>% filter(High==0) %>%
  t.test(mean_cheapest_asset ~ Tanzania, data=.)
```

In standardized units:
```{r}
df |> filter(Tanzania==1) |>
  lm(scale(mean_cheapest_asset) ~ Treatment, data=_)
df |> filter(High==0) |>
  lm(scale(mean_cheapest_asset) ~ Tanzania, data=_)
```



Correlations between risk aversion and outcomes:
```{r}
df %>% group_by(country) %>% 
  summarize(c_e1 = cor(mean_cheapest_asset, CCEI_PQR),
            c_e2 = cor(mean_cheapest_asset, FOSD_PQR),
            c_e3 = cor(mean_cheapest_asset, EU_PQR),
            c_IQ = cor(mean_cheapest_asset, iq)) %>%
  gt::gt() |>
  gt::fmt_number(columns=starts_with("c_"), decimals=3) 
```



# Main regression table: Development gap in econ rationality

```{r}
GARP1 <- lm(CCEI_PQR ~ Tanzania + High + age + female + one_parent_college + workincome + happy + yworried + government_preference, data=df)
GARP2 <- lm(CCEI_PQR ~ Tanzania + High + age + female + one_parent_college + workincome + happy + yworried + government_preference + BF_A + BF_C + BF_E + BF_N + BF_O , data=df)
GARP3 <- lm(CCEI_PQR ~ Tanzania + High + age + female + one_parent_college + workincome + happy + yworried + government_preference + BF_A + BF_C + BF_E + BF_N + BF_O + iq, data=df)

FGARP1 <- lm(FOSD_PQR ~ Tanzania + High + age + female + one_parent_college + workincome + happy + yworried + government_preference, data=df)
FGARP2 <- lm(FOSD_PQR ~ Tanzania + High + age + female + one_parent_college + workincome + happy + yworried + government_preference + BF_A + BF_C + BF_E + BF_N + BF_O , data=df)
FGARP3 <- lm(FOSD_PQR ~ Tanzania + High + age + female + one_parent_college + workincome + happy + yworried + government_preference + BF_A + BF_C + BF_E + BF_N + BF_O + iq, data=df)

EUT1 <- lm(EU_PQR ~ Tanzania + High + age + female + one_parent_college + workincome + happy + yworried + government_preference , data=df)
EUT2 <- lm(EU_PQR ~ Tanzania + High + age + female + one_parent_college + workincome + happy + yworried + government_preference + BF_A + BF_C + BF_E + BF_N + BF_O ,  data=df)
EUT3 <- lm(EU_PQR ~ Tanzania + High + age + female + one_parent_college + workincome + happy + yworried + government_preference + BF_A + BF_C + BF_E + BF_N + BF_O + iq, data=df)

IQ1 <- lm(iq ~ Tanzania + High + age + female + one_parent_college + workincome + happy + yworried + government_preference, data=df)
IQ2 <- lm(iq ~ Tanzania + High + age + female + one_parent_college + workincome + happy + yworried + government_preference + 
              BF_A + BF_C + BF_E + BF_N + BF_O, data=df)
```

Based on these regressions, I want to add the p-value of a test that there is no difference between the US and the Tanzanian high group:

```{r}
ftests <- c("US equal to Tanzania high (p-val)", 
            format(summary(glht(GARP1, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(FGARP1, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(EUT1, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(GARP2, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(FGARP2, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(EUT2, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(GARP3, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(FGARP3, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(EUT3, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(IQ1, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(IQ2, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3)
)
```

Also want to (manually) add row with standard deviation of outcomes. 
```{r}
sds <- df %>% dplyr::select(CCEI_PQR, FOSD_PQR, EU_PQR, iq) |>
  summarise_all(sd) 
sds |> gt::gt() |> gt::fmt_number(decimals=3)
```

And for the standardized effect of "Tanzania", calculate effects

```{r}
Ztanzania <- c(0,
               coef(GARP1)[['Tanzania']] / sds$CCEI_PQR[1],
               coef(FGARP1)[['Tanzania']] / sds$FOSD_PQR[1],
               coef(EUT1)[['Tanzania']] / sds$EU_PQR[1],
               coef(GARP2)[['Tanzania']] / sds$CCEI_PQR[1],
               coef(FGARP2)[['Tanzania']] / sds$FOSD_PQR[1],
               coef(EUT2)[['Tanzania']] / sds$EU_PQR[1],
               coef(GARP3)[['Tanzania']] / sds$CCEI_PQR[1],
               coef(FGARP3)[['Tanzania']] / sds$FOSD_PQR[1],
               coef(EUT3)[['Tanzania']] / sds$EU_PQR[1],
               coef(IQ1)[['Tanzania']] / sds$iq,
               coef(IQ2)[['Tanzania']] / sds$iq)
Ztanzania
```



```{r warning=FALSE}
m_names <- c('Tanzania' = 'Tanzania',
             'High' = 'High stakes (Tanzania)',
             'age' = 'Age',
             'female' = 'Female',
             'one_parent_college' = 'Parents have college education',
             'workincome' = 'Income from work?',
             'happy' = 'Happiness',
             'yworried' = 'Worried yesterday',
             'government_preferenceYes' = 'Public sector preference',
             'BF_A' = 'Agreeableness',
             'BF_C' = 'Conscientiousness',
             'BF_E' = 'Extraversion',
             'BF_N' = 'Neuroticism',
             'BF_O' = 'Openness to experience',
             'iq' = 'IQ',
             '(Intercept)' = 'Constant')
big5 <- c("Big-5 included?", "No", "No", "No", "Yes","Yes","Yes","Yes","Yes","Yes","Yes","Yes")
modelsummary::modelsummary(list("(1)"=GARP1, "(2)"=FGARP1, "(3)"=EUT1, 
                                "(4)"=GARP2, "(5)"=FGARP2, "(6)"=EUT2,
                                "(7)"=GARP3, "(8)"=FGARP3, "(9)"=EUT3,
                                "(10)"=IQ1, "(11)"=IQ2),
                           statistic = "std.error", coef_map = m_names,
                           stars =  c('*' = .1, '**' = .05, '***'=0.01),
                           gof_omit = 'R2 Adj.|AIC|BIC|Log.Lik.|F|RMSE',
                           output = 'latex',
                           add_rows = as.data.frame(t(tibble(big5, Ztanzania, ftests)))) %>%
  add_header_above(c(" " = 1, "Baseline" = 3, "Baseline + Big-5" = 3,  "Baseline + Big-5 and IQ" = 3, "Cognitive ability" = 2)) %>%
  cat(file= here::here("tables/development_gap.tex"))
```

Also outputting to html record: 
```{r warning=FALSE}
modelsummary::modelsummary(list("(1)"=GARP1, "(2)"=FGARP1, "(3)"=EUT1, 
                                "(4)"=GARP2, "(5)"=FGARP2, "(6)"=EUT2,
                                "(7)"=GARP3, "(8)"=FGARP3, "(9)"=EUT3,
                                "(10)"=IQ1, "(11)"=IQ2),
                           statistic = "std.error", coef_map = m_names,
                           stars = c('*' = .1, '**' = .05, '***'=0.01),
                           gof_omit = 'R2 Adj.|AIC|BIC|Log.Lik.|F|RMSE',
                           output = 'kableExtra',
                           add_rows = as.data.frame(t(tibble(big5, Ztanzania, ftests)))) %>%
  add_header_above(c(" " = 1, "Baseline" = 3, "Baseline + Big-5" = 3,  "Baseline + Big-5 and IQ" = 3, "Cognitive ability" = 2)) 
```






And what are the correlations between IQ and the rationalizability measurses in the US, in Tanzania low and Tanzania high stakes?

```{r}
df |> group_by(country_high) |> 
  summarize(cor_IQ_CCEI = cor(iq, CCEI_PQR),
            cor_IQ_FOSD = cor(iq, FOSD_PQR),
            cor_IQ_EU= cor(iq, EU_PQR)) |>
  gt::gt() |> 
  gt::fmt_number(decimals=3)
```




# More descriptives about the Tanzania sample


# Further analysis (for appendix)

## CDF graphs for outcomes

```{r}
tar_load(cdf_dta)
cdf_dta %>% ggplot(aes(x=value, linetype=country)) + 
  stat_ecdf(geom="step") + 
  facet_wrap(.~yfct) + 
  theme_minimal() + 
  theme(legend.position="bottom") + 
  labs(y = "ECDF(outcome)", x=element_blank()) 
ggsave(here("graphs", "cdf_outcomes.pdf"), width=16, height = 10, units = "cm")
```


## Variant of Table I from main paper with risk aversion and everything 


```{r}
rGARP1 <- lm(CCEI_PQR ~ Tanzania + High + mean_cheapest_asset + age + female + one_parent_college + workincome + happy + yworried + government_preference + gambling + funding_education , data=df)
rGARP2 <- lm(CCEI_PQR ~ Tanzania + High + mean_cheapest_asset + age + female + one_parent_college + workincome + happy + yworried + government_preference + gambling + funding_education  + BF_A + BF_C + BF_E + BF_N + BF_O , data=df)
rGARP3 <- lm(CCEI_PQR ~ Tanzania + High + mean_cheapest_asset + age + female + one_parent_college + workincome + happy + yworried + government_preference + gambling + funding_education  + BF_A + BF_C + BF_E + BF_N + BF_O + iq, data=df)

rFGARP1 <- lm(FOSD_PQR ~ Tanzania + High + mean_cheapest_asset + age + female + one_parent_college + workincome + happy + yworried + government_preference + gambling + funding_education , data=df)
rFGARP2 <- lm(FOSD_PQR ~ Tanzania + High + mean_cheapest_asset + age + female + one_parent_college + workincome + happy + yworried + government_preference + gambling + funding_education  + BF_A + BF_C + BF_E + BF_N + BF_O , data=df)
rFGARP3 <- lm(FOSD_PQR ~ Tanzania + High + mean_cheapest_asset + age + female + one_parent_college + workincome + happy + yworried + government_preference + gambling + funding_education  + BF_A + BF_C + BF_E + BF_N + BF_O + iq, data=df)

rEUT1 <- lm(EU_PQR ~ Tanzania + High + mean_cheapest_asset + age + female + one_parent_college + workincome + happy + yworried + government_preference + gambling + funding_education  , data=df)
rEUT2 <- lm(EU_PQR ~ Tanzania + High + mean_cheapest_asset + age + female + one_parent_college + workincome + happy + yworried + government_preference + gambling + funding_education  + BF_A + BF_C + BF_E + BF_N + BF_O ,  data=df)
rEUT3 <- lm(EU_PQR ~ Tanzania + High + mean_cheapest_asset + age + female + one_parent_college + workincome + happy + yworried + government_preference + gambling + funding_education  + BF_A + BF_C + BF_E + BF_N + BF_O + iq, data=df)

rIQ1 <- lm(iq ~ Tanzania + High + mean_cheapest_asset + age + female + one_parent_college + workincome + happy + yworried + government_preference + gambling + funding_education , data=df)
rIQ2 <- lm(iq ~ Tanzania + High + mean_cheapest_asset + age + female + one_parent_college + workincome + happy + yworried + government_preference + gambling + funding_education  + 
              BF_A + BF_C + BF_E + BF_N + BF_O, data=df)
```



```{r}
rm_names <- c('Tanzania' = 'Tanzania',
             'High' = 'High stakes (Tanzania)',
             'mean_cheapest_asset' = 'Risk aversion',
             'age' = 'Age',
             'female' = 'Female',
             'one_parent_college' = 'Parents have college education',
             'workincome' = 'Income from work?',
             'happy' = 'Happiness',
             'yworried' = 'Worried yesterday',
             'government_preferenceYes' = 'Public sector preference',
             'gambling' = 'Self image as gambler',
             'funding_educationYes' = 'Would spend money on education',
             'iq' = 'IQ',
             '(Intercept)' = 'Constant')
modelsummary::modelsummary(list("(1)"=rGARP1, "(2)"=rFGARP1, "(3)"=rEUT1, 
                                "(4)"=rGARP2, "(5)"=rFGARP2, "(6)"=rEUT2,
                                "(7)"=rGARP3, "(8)"=rFGARP3, "(9)"=rEUT3,
                                "(10)"=rIQ1, "(11)"=rIQ2),
                           statistic = "std.error", coef_map = rm_names,
                           stars =  c('*' = .1, '**' = .05, '***'=0.01),
                           gof_omit = 'R2 Adj.|AIC|BIC|Log.Lik.|F|RMSE',
                           output = 'latex',
                           add_rows = as.data.frame(t(big5))) %>%
  add_header_above(c(" " = 1, "Baseline" = 3, "Baseline + Big-5" = 3,  "Baseline + Big-5 and IQ" = 3, "Cognitive ability" = 2)) %>%
  cat(file= here::here("tables/development_gap_withrisk.tex"))
```

Also outputting to html record: 
```{r}
modelsummary::modelsummary(list("(1)"=rGARP1, "(2)"=rFGARP1, "(3)"=rEUT1, 
                                "(4)"=rGARP2, "(5)"=rFGARP2, "(6)"=rEUT2,
                                "(7)"=rGARP3, "(8)"=rFGARP3, "(9)"=rEUT3,
                                "(10)"=rIQ1, "(11)"=rIQ2),
                           statistic = "std.error", coef_map = rm_names,
                           stars = c('*' = .1, '**' = .05, '***'=0.01),
                           gof_omit = 'R2 Adj.|AIC|BIC|Log.Lik.|F|RMSE',
                           output = 'kableExtra',
                           add_rows = as.data.frame(t(big5))) %>%
  add_header_above(c(" " = 1, "Baseline" = 3, "Baseline + Big-5" = 3,  "Baseline + Big-5 and IQ" = 3, "Cognitive ability" = 2)) 
```

## Treatment effect of stakes on rationalizability (in Tanzania)


```{r message=FALSE, warning=FALSE, include=FALSE}
df_noUS <- df |> filter(country=="Tanzania")
tGARP1 <- lm(CCEI_PQR ~ High , data=df_noUS)
tGARP2 <- lm(CCEI_PQR ~ High + age + female + one_parent_college + workincome + happy + yworried + government_preference + BF_A + BF_C + BF_E + BF_N + BF_O , data=df_noUS)
tGARP3 <- lm(CCEI_PQR ~ High + age + female + one_parent_college + workincome + happy + yworried + government_preference + BF_A + BF_C + BF_E + BF_N + BF_O +iq, data=df_noUS)

tFGARP1 <- lm(FOSD_PQR ~ High , data=df_noUS)
tFGARP2 <- lm(FOSD_PQR ~ High + age + female + one_parent_college + workincome + happy + yworried + government_preference + BF_A + BF_C + BF_E + BF_N + BF_O , data=df_noUS)
tFGARP3 <- lm(FOSD_PQR ~ High + age + female + one_parent_college + workincome + happy + yworried + government_preference + BF_A + BF_C + BF_E + BF_N + BF_O + iq, data=df_noUS)

tEUT1 <- lm(EU_PQR ~ High , data=df_noUS)
tEUT2 <- lm(EU_PQR ~ High + age + female + one_parent_college + workincome + happy + yworried + government_preference + BF_A + BF_C + BF_E + BF_N + BF_O , data=df_noUS)
tEUT3 <- lm(EU_PQR ~ High + age + female + one_parent_college + workincome + happy + yworried + government_preference + BF_A + BF_C + BF_E + BF_N + BF_O + iq, data=df_noUS)
```


```{r}
t_names <- c('High' = 'High stakes',
             'mean_cheapest_asset' = 'Risk aversion',
             'age' = 'Age',
             'female' = 'Female',
             'one_parent_college' = 'Parents have college education',
             'workincome' = 'Income from work?',
             'happy' = 'Happiness',
             'yworried' = 'Worried yesterday',
             'government_preferenceYes' = 'Public sector preference',
             'iq' = 'IQ',
             '(Intercept)' = 'Constant')
t_big5 <- c("Big-5 included?", "No", "No", "No", "Yes","Yes","Yes", "Yes","Yes","Yes")
modelsummary::modelsummary(list("(1)"=tGARP1, "(2)"=tFGARP1, "(3)"=tEUT1, 
                                "(4)"=tGARP2, "(5)"=tFGARP2, "(6)"=tEUT2,
                                "(7)"=tGARP3, "(8)"=tFGARP3, "(9)"=tEUT3),
                           statistic = "std.error", coef_map = t_names,
                           stars = c('*' = .1, '**' = .05, '***'=0.01),
                           gof_omit = 'R2 Adj.|AIC|BIC|Log.Lik.|F|RMSE',
                           output = 'latex', 
                           add_rows = as.data.frame(t(t_big5))) |>
  add_header_above(c(" " = 1, "No controls" = 3, "Baseline controls w. Big-5" = 3,  "Baseline controls w. Big-5 and IQ" = 3)) %>%
  cat(file= here::here("tables/treatmenteffect_stakes.tex"))
```

Also outputting to html record: 
```{r}
modelsummary::modelsummary(list("(1)"=tGARP1, "(2)"=tFGARP1, "(3)"=tEUT1, 
                                "(4)"=tGARP2, "(5)"=tFGARP2, "(6)"=tEUT2,
                                "(7)"=tGARP3, "(8)"=tFGARP3, "(9)"=tEUT3),
                           statistic = "std.error", coef_map = t_names,
                           stars = c('*' = .1, '**' = .05, '***'=0.01),
                           gof_omit = 'R2 Adj.|AIC|BIC|Log.Lik.|F|RMSE',
                           output = 'kableExtra', 
                           add_rows = as.data.frame(t(t_big5))) %>%
  add_header_above(c(" " = 1, "No controls" = 3, "Baseline controls w. Big-5" = 3,  "Baseline controls w. Big-5 and IQ" = 3)) 
```


## Treatment effect of (high) stakes on risk aversion (in Tanzania)

```{r message=FALSE, warning=FALSE, include=FALSE}
r1 <- lm(mean_cheapest_asset ~ High , data=df_noUS)
r2 <- lm(mean_cheapest_asset ~ High + age + female + one_parent_college + workincome + happy + yworried + government_preference + BF_A + BF_C + BF_E + BF_N + BF_O , data=df_noUS)
r3 <- lm(mean_cheapest_asset ~ High + age + female + one_parent_college + workincome + happy + yworried + government_preference + BF_A + BF_C + BF_E + BF_N + BF_O + iq, data=df_noUS)
```


```{r}
r_names <- c('High' = 'High stakes',
             'mean_cheapest_asset' = 'Risk aversion',
             'age' = 'Age',
             'female' = 'Female',
             'one_parent_college' = 'Parents have college education',
             'workincome' = 'Income from work?',
             'happy' = 'Happiness',
             'yworried' = 'Worried yesterday',
             'government_preferenceYes' = 'Public sector preference',
             'iq' = 'IQ',
             '(Intercept)' = 'Constant')


r_big5 <- c("Big-5 included?", "No","Yes","Yes")
modelsummary::modelsummary(list("(1)"=r1, "(2)"=r2, "(3)"=r3),
                           statistic = "std.error", coef_map = r_names,
                           stars = c('*' = .1, '**' = .05, '***'=0.01),
                           gof_omit = 'R2 Adj.|AIC|BIC|Log.Lik.|F|RMSE',
                           add_rows = as.data.frame(t(r_big5)),
                           output = 'latex') |>
  add_header_above(c(" ","Mean share cheapest asset"=3)) |>
  cat(file= here::here("tables/treatmenteffect_riskaversion.tex"))
```

Also outputting to html record: 
```{r}
modelsummary::modelsummary(list("(1)"=r1, "(2)"=r2, "(3)"=r3),
                           statistic = "std.error", coef_map = r_names,
                           stars = c('*' = .1, '**' = .05, '***'=0.1),
                           gof_omit = 'R2 Adj.|AIC|BIC|Log.Lik.|F|RMSE',
                           add_rows = as.data.frame(t(r_big5)),
                           output = 'kableExtra') %>%
  add_header_above(c(" ","Mean share cheapest asset"=3))
```

## Distribution of happiness

First, the averages:
```{r}
df |> group_by(country) |> 
  summarize(mean_happiness = mean(happy, na.rm=TRUE)) |>
  gt::gt() |> gt::fmt_number(decimals=3)
```


```{r}
df |> group_by(country, happy) |> 
  summarize( freq = n()) |>
  group_by(country) |>
  mutate(proportion = freq / sum(freq)) |>
  ggplot(aes(x=happy, y = proportion, fill=country)) + 
  geom_bar(stat='identity', position = position_dodge(preserve = "single"), width=0.9, preserve='single') + 
  labs(x = "Self-reported happiness", y="Fraction") + 
  scale_x_continuous(breaks = c(0,5,10)) +
  theme_minimal()
ggsave(here::here("graphs","happiness_distribution.pdf"), width=16, height=10, units="cm")
```

## Breaking down by preference for public sector

The question asked was 

> Assume that you could choose between the following job opportunities, and that income and work hours were exactly the
same in all of them. How would you rate these jobs?

They then rated "government employee" on a 1 (not so good) to 3 (very good) scale. Breaking down such that 3 is
an active preference from being a government employee:

```{r}
tar_load(pubpreffig_dta)
pubpreffig_dta |>
  ggplot(aes(x=outcome_nm, 
             y=mean, 
             ymin = mean - 1.96 * se_mean, 
             ymax = mean + 1.96 * se_mean, 
             fill=government_preference)) + 
  geom_bar(orientation="x", stat="identity", position=position_dodge2()) +
  geom_errorbar(position=position_dodge(0.9), width=0.3) +
  labs(y="Mean \u00B1 95% C.I.", x= element_blank(), fill = "Government\npreference?") +
  theme_minimal() + 
  ylim(0, 1) + 
  facet_wrap(. ~ country)

ggsave(here("graphs", "government_preference.pdf"), width=16, height=10, units="cm")

```




# Session Info
```{r}
sessionInfo()
```

