---
title: "The Development Gap in Economic Rationality of Future Elites"
author: "Erik Ø. Sørensen"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 2
  pdf_document:
    toc: yes
    toc_depth: '2'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
extrafont::loadfonts()
library("multcomp")
library("tidyverse")
library("haven")
library("modelsummary")
library("kableExtra")
library("here")
```


# Reading data
Reading the background data
```{r}
dfb <- read_dta(here("data", "background.dta")) %>% 
  mutate(two_parents_HS = mother2school*father2school,
         one_parent_college = pmax(motheruni,fatheruni),
         country = sjlabelled::as_label(country),
         iq = iq/26.0,
         female = as.numeric(sex==2),
         Tanzania = as.numeric( country=="Tanzania"))
```

Reading the rationalizability scores.
```{r}
df <- read_dta(here("data", "calculated_RP_measures.dta")) |> 
  left_join(dfb)
```


# Summaries of background data

```{r}
dfb %>% group_by(country) %>%
  summarize(one_parent_college = mean(one_parent_college),
            female = mean(sex==2),
            age = mean(age, na.rm=TRUE)) %>%
  kable(digits=c(0,3,3,1))
```


# Graphs of behavioral outcomes

## Means of the main outcomes

Overall means for a striking first figure of the rationalizability scores. First, calculate the means and standard errors 
for both the original and the standardized scores.

```{r}
fig_dta <- df %>% 
  dplyr::select(country, CCEI_PQR, FOSD_PQR, EU_PQR, iq) %>%
  gather(key="outcome", value="value", CCEI_PQR:iq, factor_key=TRUE) %>%
  mutate(outcome_nm = fct_recode(outcome,
                                 "e*" = "CCEI_PQR",
                                 "e**" = "FOSD_PQR",
                                 "e***" = "EU_PQR"))  %>%
  group_by(outcome_nm) %>%
  mutate(zvalue = scale(value),
         sd = sd(value)) %>%
  group_by(country, outcome_nm) %>%
  summarize(overall_sd = min(sd),
            mean = mean(value),
            zmean = mean(zvalue),
            se_mean = sd(value)/sqrt(n()),
            se_zmean = sd(zvalue)/sqrt(n()))
```

Now, the first plot with the rationalizability scores:
```{r}
fig_dta %>% filter(outcome_nm!="iq") %>%
  ggplot(aes(x=outcome_nm, 
             y=mean, 
             ymin = mean - 1.96 * se_mean, 
             ymax = mean + 1.96 * se_mean, 
             fill=country)) + 
  geom_bar(orientation="x", stat="identity", position=position_dodge2()) +
  geom_errorbar(position=position_dodge(0.9), width=0.3) +
  labs(y="Mean \u00B1 95% C.I.", x= element_blank(), fill = "Country") +
  theme_minimal() + 
  ylim(0, 1) 
ggsave(here("graphs", "mean_differences.pdf"), width=16, height=10, units="cm")
```

Similar graph, but with standardized outcomes and including the IQ scores. We decided to change this
graph into differences between countries: I make it differences between US and Tanzania.

```{r}
fig_dtaw <- fig_dta |> 
  pivot_wider(id_cols = outcome_nm, names_from = country, values_from = c(zmean, se_zmean)) |>
  mutate( zdiff = `zmean_United States` - zmean_Tanzania,
          zdiff_se =  sqrt((`se_zmean_United States`)^2 + se_zmean_Tanzania^2 ))
```


```{r}
fig_dtaw %>%
  ggplot(aes(x=outcome_nm, 
             y=zdiff, 
             ymin = zdiff - 1.96 * zdiff_se, 
             ymax = zdiff + 1.96 * zdiff_se)) + 
  geom_col() +
  geom_errorbar(width=0.3) +
  labs(y="Mean difference in standardized outcomes\nUnited States vs Tanzania \u00B1 95% C.I.", x= element_blank()) +
  theme_minimal() 
ggsave(here("graphs", "zmean_differences.pdf"), width=16, height=10, units="cm")
```

In the final figure, each of the four outcomes have been standardized to an overall average of zero and unit variance before
calculating the averages by country.

What are the numbers?
```{r}
fig_dta %>% dplyr::select(country, outcome_nm, overall_sd, mean, se_mean, zmean, se_zmean) %>%
  knitr::kable(digits=3)
```


## CDF graphs for all outcomes

For reference, not for the paper: CDFs for the four outcomes on the original scale.
```{r}
cdf_data <- df %>% 
  dplyr::select(country, CCEI_PQR, FOSD_PQR, EU_PQR, iq) %>%
  gather(key="outcome", value="value", CCEI_PQR:iq, factor_key=TRUE) %>%
  filter(!is.na(country)) %>%
  mutate(yfct=fct_recode(outcome,
                      "e*" = "CCEI_PQR",
                      "e**" = "FOSD_PQR",
                      "e***" = "EU_PQR"))
cdf_data %>% ggplot(aes(x=value, linetype=country)) + 
  stat_ecdf(geom="step") + facet_wrap(.~yfct) + 
  theme_minimal() + theme(legend.position="bottom") + 
  ylab("ECDF(outcome)")  +
  xlab(element_blank()) 
ggsave(here("graphs", "cdf_outcomes.pdf"))
```

## Survival function plot 
```{r}
s_df <- cdf_data %>% group_by(country,yfct) %>% 
  mutate(S  = 1 - cume_dist(value),
         prop = cut(value, c(0,seq(0.6,1,0.1)))) %>%
  group_by(country, yfct, prop) %>%
  summarize(meanS = mean(S),
            maxS = max(S))

s_df %>% filter(prop != "(0,0.6]") %>%
  filter(yfct !="IQ") %>%
  mutate(interval_lower = fct_recode(prop,
                                     "0.6" = "(0.6,0.7]",
                                     "0.7" = "(0.7,0.8]",
                                     "0.8" = "(0.8,0.9]",
                                     "0.9" = "(0.9,1]"),
         yfct = fct_recode(yfct,
                           "A.           e*" = "e*",
                           "B.           e**"= "e**",
                           "C.           e***"="e***")) %>%
  ggplot(aes(y=maxS, x=interval_lower, group=country, fill=country)) +
  geom_bar(stat='identity', position = position_dodge2()) +
  theme_minimal() +
  facet_wrap(. ~ yfct) + 
  theme( strip.text=element_text(hjust=0) ) +
  labs(x = "Critical value",
       y = "Fraction of subjects") 
ggsave(here::here("graphs","survival_graph.pdf"), 
       width=16, height=10, units="cm")
```

For reference, but not for the paper: Continuous variants of the survival graphs for all four outcomes:
```{r}
cdf_data %>% group_by(country,yfct) %>% 
  filter(value>0.6) %>%
  mutate(S  = 1 - cume_dist(value)) %>%
  ggplot(aes(y = S, x = value, group=country, linetype=country)) +
  geom_step() +
  theme_minimal() +
  facet_wrap(.~yfct) +
  labs(x = "Efficiency index",
       y = "Proportion > efficiency index",
       linetype = "Country")
ggsave(here::here("graphs","survival_graph_continuous.pdf"),
       width=16, height = 10, units="cm")
```


# Summary statistics on socio demographics


```{r}
dfb_selected <- dfb %>% dplyr::select(c("ID", "workincome", "two_parents_HS", 
                                 "one_parent_college"))
se <- function(x) {
  sd(x, na.rm=TRUE)/sqrt(length(x[!is.na(x)]))
}
df %>% dplyr::select(c("ID", "country", "age", "female", "BF_E", "BF_A", "BF_C", 
  "BF_N", "BF_O", "iq")) %>%
  left_join(dfb_selected, by="ID") %>%
  dplyr::select(-"ID") %>%
  group_by(country) %>%
  summarize_all(list(mean= ~ mean(., na.rm=TRUE), se= ~se(.)))  %>%
  gather(key="key", value="value", Age_mean:one_parent_college_se) %>%
  spread(key="country", value="value") %>% knitr::kable(digits=3)
```


## Tests for table
```{r}
df %>% t.test(iq ~ country, data=.)
df %>% t.test(BF_A ~ country, data=.)
df %>% t.test(BF_C ~ country, data=.)
df %>% t.test(BF_E ~ country, data=.)
df %>% t.test(BF_N ~ country, data=.)
df %>% t.test(BF_O ~ country, data=.)
```


# Risk aversion

For Tanzania, we are interested in whether high and low stakes individuals
exhibited different risk aversion. We operationalize risk aversion by
the average fraction of tokens in the cheaper security.

First, read the decision data, drop the pilot data, 
calculate the mean share to the cheaper asset for everyone, join this onto 
the other data at the individual level.
```{r}
decisions <- read_dta(here("data", "decisions.dta")) 
ra_df <- decisions %>% 
  filter(ID>100) %>%
  mutate(share_cheapest = ifelse(YM>XM, Y/(X+Y), X/(X+Y))) %>%
  group_by(ID, Treatment) %>%
  summarize(mean_cheapest_asset = mean(share_cheapest))
df <- df %>% left_join(ra_df)
```

Now, what are the average risk aversion in the three group?
```{r}
dfsum <- df %>% mutate(country_high = factor(ifelse(Tanzania!=1, "United States", 
                                      ifelse(High==1, "Tanzania - high", "Tanzania - low"))))

dfsum %>% group_by(country_high) %>% summarize(mean = mean(mean_cheapest_asset),
                                       std.dev = sd(mean_cheapest_asset),
                                       se = std.dev/sqrt(n())) %>%
  knitr::kable(digits=3)
```

Also, the distribution of this measure across groups:
```{r}
dfsum %>% ggplot(aes(x=mean_cheapest_asset, 
                     y=(..count..)/tapply(..count..,..PANEL..,sum)[..PANEL..])) + 
  geom_histogram( binwidth = 0.05, boundary=0.5) + 
  theme_minimal() +
  facet_wrap(. ~ country_high) +
  labs(y = "Proportion of individuals",
       x = "Mean share of tokens allocated to cheapest asset")
```


Testing (t-test) for equality between the treatments in Tanzania?
```{r}
df %>% filter(Tanzania==1) %>%
  t.test(mean_cheapest_asset ~ High, data=.)
```

Correlations between risk aversion and outcomes:
```{r}
df %>% group_by(Country) %>% 
  summarize(c_e1 = cor(mean_cheapest_asset, CCEI_PQR),
            c_e2 = cor(mean_cheapest_asset, FOSD_PQR),
            c_e3 = cor(mean_cheapest_asset, EU_PQR),
            c_IQ = cor(mean_cheapest_asset, IQ)) %>%
  knitr::kable(digits=3)
```



# Development gap in econ rationality

Adding the background data (for parental education).
```{r}
df_wb <- df %>% left_join(dfb_selected, by="ID")
```


```{r}
GARP1 <- lm(CCEI_PQR ~ Tanzania + High + mean_cheapest_asset + Age + Female + one_parent_college, data=df_wb)
GARP2 <- lm(CCEI_PQR ~ Tanzania + High + mean_cheapest_asset + Age + Female + one_parent_college + BF_A + BF_C + BF_E + BF_N + BF_O , data=df_wb)
GARP3 <- lm(CCEI_PQR ~ Tanzania + High + mean_cheapest_asset + Age + Female + one_parent_college + BF_A + BF_C + BF_E + BF_N + BF_O + IQ, data=df_wb)

FGARP1 <- lm(FOSD_PQR ~ Tanzania + High + mean_cheapest_asset + Age + Female + one_parent_college, data=df_wb)
FGARP2 <- lm(FOSD_PQR ~ Tanzania + High + mean_cheapest_asset + Age + Female + one_parent_college + BF_A + BF_C + BF_E + BF_N + BF_O , data=df_wb)
FGARP3 <- lm(FOSD_PQR ~ Tanzania + High + mean_cheapest_asset + Age + Female + one_parent_college + BF_A + BF_C + BF_E + BF_N + BF_O + IQ, data=df_wb)

EUT1 <- lm(EU_PQR ~ Tanzania + High + mean_cheapest_asset + Age + Female + one_parent_college , data=df_wb)
EUT2 <- lm(EU_PQR ~ Tanzania + High + mean_cheapest_asset + Age + Female + one_parent_college + BF_A + BF_C + BF_E + BF_N + BF_O ,  data=df_wb)
EUT3 <- lm(EU_PQR ~ Tanzania + High + mean_cheapest_asset + Age + Female + one_parent_college + BF_A + BF_C + BF_E + BF_N + BF_O + IQ, data=df_wb)

IQ1 <- lm(IQ ~ Tanzania + High + mean_cheapest_asset + Age + Female + one_parent_college, data=df_wb)
IQ2 <- lm(IQ ~ Tanzania + High + mean_cheapest_asset + Age + Female + one_parent_college + 
              BF_A + BF_C + BF_E + BF_N + BF_O, data=df_wb)
```

Based on these regressions, I want to add the p-value of a test that there is no difference between the US and the Tanzanian high group:

```{r}
ftests <- c("US equal to Tanzania high (p-val)", 
            format(summary(glht(GARP1, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(FGARP1, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(EUT1, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(GARP2, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(FGARP2, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(EUT2, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(GARP3, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(FGARP3, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(EUT3, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(IQ1, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3),
            format(summary(glht(IQ2, linfct = c("Tanzania + High = 0")))$test$pvalues[1],digits=3)
)
```



```{r}
m_names <- c('Tanzania' = 'Tanzania',
             'High' = 'High stakes (Tanzania)',
             'mean_cheapest_asset' = 'Risk aversion',
             'Age' = 'Age',
             'Female' = 'Female',
             'one_parent_college' = 'Parents have college education',
             'IQ' = 'IQ')

modelsummary::modelsummary(list("(1)"=GARP1, "(2)"=FGARP1, "(3)"=EUT1, 
                                "(4)"=GARP2, "(5)"=FGARP2, "(6)"=EUT2,
                                "(7)"=GARP3, "(8)"=FGARP3, "(9)"=EUT3,
                                "(10)"=IQ1, "(11)"=IQ2),
                           statistic = "std.error", coef_map = m_names,
                           stars = TRUE,
                           gof_omit = 'R2 Adj.|AIC|BIC|Log.Lik.|F',
                           output = 'latex',
                           add_rows = as.data.frame(t(ftests))) %>%
  add_header_above(c(" " = 1, "Baseline" = 3, "Baseline + Big-5" = 3,  "Baseline + Big-5 and IQ" = 3, "Cognitive ability" = 2)) %>%
  cat(file= here::here("development_gap.tex"))
```

Also outputting to html record: 
```{r}
modelsummary::modelsummary(list("(1)"=GARP1, "(2)"=FGARP1, "(3)"=EUT1, 
                                "(4)"=GARP2, "(5)"=FGARP2, "(6)"=EUT2,
                                "(7)"=GARP3, "(8)"=FGARP3, "(9)"=EUT3,
                                "(10)"=IQ1, "(11)"=IQ2),
                           statistic = "std.error", coef_map = m_names,
                           stars = TRUE,
                           gof_omit = 'R2 Adj.|AIC|BIC|Log.Lik.|F',
                           output = 'kableExtra',
                           add_rows = as.data.frame(t(ftests))) %>%
  add_header_above(c(" " = 1, "Baseline" = 3, "Baseline + Big-5" = 3,  "Baseline + Big-5 and IQ" = 3, "Cognitive ability" = 2)) 
```



Also want to add row with standard deviation of outcomes. 
```{r}
df %>% dplyr::select(CCEI_PQR, FOSD_PQR, EU_PQR, IQ) %>%
  summarise_all(sd) %>%
  knitr::kable(digits=3)
```




# More descriptives about the samples

For the Tanzania sample, we have a separate survey. 
```{r}
tz_df <- read_dta(here::here("data","tanzaniasurvey.dta"))
```

Many of the answers we are interested in are free-form, such as field of study and
place of origin. I dump these to csv-files, for manual categorization in a spreadsheet,
for later to be read as categories. 
```{r}
tz_df %>% group_by(birthplace) %>%
  summarize(n_responses = n()) %>%
  write_csv(here::here("birthplaces_raw.csv"))
tz_df %>% group_by(studysubject) %>%
  summarize(n_responses = n()) %>%
  write_csv(here::here("studysubjects_raw.csv"))
birthplaces_categorized <- read_csv(here::here("birthplaces_categorized.csv")) %>%
  dplyr::select(-n_responses)
studysubjects_categorized <- read_csv(here::here("studysubjects_categorized.csv")) %>%
  dplyr::select(-n_responses)
tz_df <- tz_df %>% left_join(studysubjects_categorized, by="studysubject") %>%
  left_join(birthplaces_categorized, by="birthplace")
```

## Where are they from? 
There are 32 **regions** in Tanzania (three of them on Zanzibar). I have 
manually classified self-reported birthplace to region.

```{r}
tz_df %>% group_by(region) %>%
  summarize( n = n()) %>%
  arrange(-n) %>% 
  mutate( proportion = n/sum(n), 
          ranking = row_number()) %>%
  knitr::kable(digits=3)
```


Note that the one "Zanzibar" response is not allocated to a proper region, as
no more information than "Zanzibar" was given by the respondent. 

While the participants at lab experiments at major American universities
have been extensively studied, our sample at the University of Dar-es-Salaam
comes from a less studied population, and not from a permanent pool
of lab experiments. We implemented a small survey of our Tanzanian
lab participants (on 215 out of our 216 lab participants). 

While the university is located in Dar-es-Salaam, we have participants
from 25 out of the 30 Tanzanian regions (as of 2012). Those born in 
Dar-es-Salaam are well represented (16.7 %), but there are slightly more 
from Kagera (17.2%) and almost as many from Kilimanjaro (13.5 %). There 
are 6% or less from each of the other regions.

```{r}
tz_df %>% group_by(category) %>%
  summarize( n = n()) %>%
  arrange(-n) %>% 
  mutate( proportion = n/sum(n), 
          ranking = row_number()) %>%
  knitr::kable(digits=3)
```
The most popular reported field of study among our respondents is one or more of the social sciences (42.8%),
with science and engineering close behind (31.6%). Education (9.8%) and humanities (8.4%) are also represented,
but there are less than four percent each from business management and law. On average they report
to be `r mean(tz_df$yrs_study, na.rm=TRUE)` years into their study, with `r max(tz_df$yrs_study, na.rm=TRUE)` as the maximum. 

The median reported yearly expenses was `r median(tz_df$year_expenses2, na.rm=TRUE)` TZS, corresponding to about 
`r 0.0006 * median(tz_df$year_expenses2, na.rm=TRUE)` USD at 
2012 exchange rates. Many report that they have support from their family (`r 100*mean(tz_df$familysupport)`%), 
with a median support of `r median(tz_df$familysupport_value, na.rm=TRUE)` TZS per week (`r 0.0006 * median(tz_df$familysupport_value, na.rm=TRUE)` USD) per 
week, and `r sum(!is.na(tz_df$govloan))/nrow(tz_df)` report that they are at least partly funded by 
government loans (with, on average, `r mean(tz_df$govloan, na.rm=TRUE)`% of the costs of study covered by the student loans).

# Session Info
```{r}
sessionInfo()
```

